A Descriptive Analysis of How Psychology Journals Handle Post-Publication Critiques

### 2024-01-03 manuscript_NHB.Rmd
Added CIs in code
Updated to calculate CIs and include in-text through code

```{r setup, include = FALSE}
renv::activate()

library(papaja) # article template
library(tidyverse)
library(ggplot2) # figure creation
library(ggrepel) # for ggplot text labels
library(patchwork) # for multipanel plot layouts
library(RColorBrewer) # data visualisation
library(dplyr)
library(kableExtra) # for tables
library(knitr) # for literate programming
library(janitor) # for data munging
library(tidylog) # for inline code feedback
library(here) # for finding files
```

```{r echo=FALSE, message=FALSE}
# Load processed data files

d_articles <- read_csv(here('ProjCode', 'GitClone', 'PROJ_2023_psych-journal-critique', 'data','processed','d_articles.csv'), show_col_types = F) # article data 
d_journals <- read_csv(here('ProjCode', 'GitClone', 'PROJ_2023_psych-journal-critique', 'data','processed','d_journals.csv'), show_col_types = F) # journal data 
```

@@

# Create the second graph
graph2 <- dj_prominent %>%
  filter(ppc_type == "A") %>% # Keeps only first entry for each journal
  filter(!is.na(has_ppc)) %>% # Remove journal rows with no (NA) PPC options 
  mutate(has_ppc = case_when(
    has_ppc == 'ARCHIVE' ~ 'Archive',
    has_ppc == 'NO EXPLICIT' ~ 'No Explicit',
    has_ppc == 'NO IMPLICIT' ~ 'No Implicit',
    has_ppc == 'YES' ~ 'Yes',
    TRUE ~ has_ppc
  )) %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = "", y = count, fill = has_ppc)) +
  geom_bar(stat = "identity",
           colour = "black") +
  theme_bw() +
  scale_fill_brewer(palette = "Oranges") +
  labs(x = "Journal Policy", y = "Count", fill = "Accepts PPC")

@@

@@

  As shown in Figure *XX*, post-publication critique policies were explicitly stated by `r dj_prominent %>% filter(ppc_type == "A") %>% summarise(percentage = sum(has_ppc == "YES") / n() * 100)` of journals, and we inferred that they were implicitly accepted (as denoted by archival examples) in `r dj_prominent %>% filter(ppc_type == "A") %>% summarise(percentage = sum(has_ppc == "ARCHIVE") / n() * 100)` of journals. `r dj_prominent %>% filter(ppc_type == "A") %>% summarise(percentage = sum(has_ppc == "NO IMPLICIT") / n() * 100)` had no advertised post-publication critique policy, and `r dj_prominent %>% filter(ppc_type == "A") %>% summarise(percentage = sum(has_ppc == "NO EXPLICIT") / n() * 100)` of journals explicitly stated that they did not accept critiques (*Clinical Psychology: Science and Practice and Psychological Bulletin*). `r dj_prominent %>% filter(ppc_type == "B" & !is.na(ppc_name)) %>% summarise(percentage = sum(!is.na(ppc_name) / n() * 100))` of journals accepted two types (commentaries and letters); *International Journal of Mental Health and Addiction*, *Journal of Environmental Psychology*, and *Psychological Science*. 
    Among the 41 explicitly stated post-publication critique options, restrictions on length were quantitatively stated (e.g., '1,000 words') in `r print(paste(prom_quantitative_length$count, "(", round(prom_quantitative_length$percentage), "%)", sep = ""))` cases, qualitatively stated (e.g. 'brief') in  `r print(paste(prom_qualitative_length$count, "(", round(prom_qualitative_length$percentage), "%)", sep = ""))` cases, and not stated in `r print(paste(prom_notStated_length$count, "(", round(prom_notStated_length$percentage), "%)", sep = ""))` cases. Time-to-submit was quantitatively stated (e.g., '4 weeks') in `r print(paste(prom_quantitative_time$count, "(", round(prom_quantitative_time$percentage), "%)", sep = ""))` cases, qualitatively stated (e.g., 'timely') in `r print(paste(prom_qualitative_time$count, "(", round(prom_qualitative_time$percentage), "%)", sep = ""))` cases, and not stated in `r print(paste(prom_notStated_time$count, "(", round(prom_notStated_time$percentage), "%)", sep = ""))` cases. Number of references allowed was quantitatively stated (e.g., 'max 20') in `r print(paste(prom_quantitative_ref$count, "(", round(prom_quantitative_ref$percentage), "%)", sep = ""))` cases and not stated in `r print(paste(prom_notStated_ref$count, "(", round(prom_notStated_ref$percentage), "%)", sep = ""))`. Whether critiques were sent for external peer-review was only stated in `r print(paste((prom_yes_review$count + prom_no_review$count), "(", round(prom_yes_review$percentage + prom_no_review$percentage), "%)", sep = ""))` cases, in `r print(paste(prom_yes_review$count, "(", round(prom_yes_review$percentage), "%)", sep = ""))` cases this was in the affirmative. 
    When stated quantitatively, length limits ranged from `r prom_length_range %>% pull(prom_min_length)` - `r prom_length_range %>% pull(prom_max_length)` words, time-to-submit limits ranged from `r prom_time_range %>% pull(prom_min_time)` -`r prom_time_range %>% pull(prom_max_time)` weeks, reference limits ranged between `r prom_ref_range %>% pull(prom_min_ref)` -`r prom_ref_range %>% pull(prom_max_ref)`. Table *XX* shows descriptive statistics for stated quantitative restrictions, broken down by critique type.

@@

# Create the second graph
graph3 <- d_journals %>%
  filter(sample_id == "random" | sample_id == "both") %>%
  filter(ppc_type == "A") %>%
  filter(!is.na(has_ppc)) %>%
  mutate(has_ppc = case_when(
    has_ppc == 'ARCHIVE' ~ 'Archive',
    has_ppc == 'NO EXPLICIT' ~ 'No Explicit',
    has_ppc == 'NO IMPLICIT' ~ 'No Implicit',
    has_ppc == 'YES' ~ 'Yes',
    TRUE ~ has_ppc
  )) %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = "", y = count, fill = has_ppc)) +
  geom_bar(stat = "identity",
           colour = "black") +
  theme_bw() +
  scale_fill_brewer(palette = "Oranges") +
  labs(x = "Journal Policy", y = "Count", fill = "Accepts PPC")

# Arrange both graphs in the same figure
combined_plots <- graph3 + graph4
combined_plots

ggsave(filename = "Random_statements.png", plot = last_plot(), device = "png")

@@

```
  As shown in Figure *XX*, post-publication critique policies were explicitly stated by `r dj_random %>% filter(ppc_type == "A") %>% summarise(percentage = sum(has_ppc == "YES") / n() * 100)`% of journals (95% CI [15.8%, 32.2%]), and we inferred that they were implicitly accepted (as denoted by archival examples) in `r dj_random %>% filter(ppc_type == "A") %>% summarise(percentage = sum(has_ppc == "ARCHIVE") / n() * 100)`% of journals (95% CI [1.0%, 8.5%]). `r dj_random %>% filter(ppc_type == "A") %>% summarise(percentage = sum(has_ppc == "NO IMPLICIT") / n() * 100)`% of journals (95% CI [63.6%, 80.7%]) had no advertised post-publication critique policy, and `r dj_random %>% filter(ppc_type == "A") %>% summarise(percentage = sum(has_ppc == "NO EXPLICIT") / n() * 100)` of journals (95% CI [0.2%, 5.4%]) explicitly stated that they did not accept critiques (*Clinical Psychology: Science and Practice and Psychological Bulletin*, also included in sample for prominent journals). `r dj_random %>% filter(ppc_type == "B" & !is.na(ppc_name)) %>% summarise(percentage = sum(!is.na(ppc_name) / n() * 100))` of journals accepted two types (95% CI [0.0%, 7.0%]); *Cortex* offered commentaries and 'verification reports' (Chambers, 2020) and *Journal of Cognition and Culture* offered commentaries and letters.

**ABOVE EDIT: insert code for CIs**

>>>>

```{r setup, include = FALSE}
#renv::activate()

library(papaja) # article template
library(tidyverse)
library(ggplot2) # figure creation
library(ggrepel) # for ggplot text labels
library(patchwork) # for multipanel plot layouts
library(RColorBrewer) # data visualisation
library(dplyr)
library(kableExtra) # for tables
library(knitr) # for literate programming
library(janitor) # for data munging
library(tidylog) # for inline code feedback
library(here) # for finding files
library(binom) # for calculating CIs
```

```{r echo=FALSE, message=FALSE}
# Load processed data files
#'ProjCode', 'GitClone', 'PROJ_2023_psych-journal-critique',
d_articles <- read_csv(here('data','processed','d_articles.csv'), show_col_types = F) # article data 
d_journals <- read_csv(here('data','processed','d_journals.csv'), show_col_types = F) # journal data 
```

@@

# Create the second graph
graph2 <- dj_prominent %>%
  filter(ppc_type == "A") %>% # Keeps only first entry for each journal
  filter(!is.na(has_ppc)) %>% # Remove journal rows with no (NA) PPC options 
  mutate(has_ppc = case_when(
    has_ppc == 'ARCHIVE' ~ 'Yes Implicit',
    has_ppc == 'NO EXPLICIT' ~ 'No Explicit',
    has_ppc == 'NO IMPLICIT' ~ 'No Implicit',
    has_ppc == 'YES' ~ 'Yes Explicit',
    TRUE ~ has_ppc
  )) %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = "", y = count, fill = has_ppc)) +
  geom_bar(stat = "identity",
           colour = "black") +
  theme_bw() +
  scale_fill_brewer(palette = "Oranges") +
  labs(x = "Journal Policy", y = "Count", fill = "Accepts PPC")

@@

# Calculate the count and percentage for 2 types
prominent2types <- dj_prominent %>%
  filter(ppc_type == "B") %>%
  filter(!is.na(ppc_name)) %>%
  count()

@@

  As shown in Figure *XX*, post-publication critique policies were explicitly stated by `r dj_prominent %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "YES") / n() * 100, "%"))` of journals, and we inferred that they were implicitly accepted (as denoted by archival examples) in `r dj_prominent %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "ARCHIVE") / n() * 100, "%"))` of journals. `r dj_prominent %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "NO IMPLICIT") / n() * 100, "%"))` had no advertised post-publication critique policy, and `r dj_prominent %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "NO EXPLICIT") / n() * 100, "%"))` of journals explicitly stated that they did not accept critiques (*Clinical Psychology: Science and Practice and Psychological Bulletin*). `r paste0(prominent2types, "%")` of journals accepted two types (commentaries and letters); *International Journal of Mental Health and Addiction*, *Journal of Environmental Psychology*, and *Psychological Science*. 
    Among the 41 explicitly stated post-publication critique options, restrictions on length were quantitatively stated (e.g., '1,000 words') in `r print(paste(prom_quantitative_length$count, "(", round(prom_quantitative_length$percentage), "%)", sep = ""))` cases, qualitatively stated (e.g. 'brief') in  `r print(paste(prom_qualitative_length$count, "(", round(prom_qualitative_length$percentage), "%)", sep = ""))` cases, and not stated in `r print(paste(prom_notStated_length$count, "(", round(prom_notStated_length$percentage), "%)", sep = ""))` cases. Time-to-submit was quantitatively stated (e.g., '4 weeks') in `r print(paste(prom_quantitative_time$count, "(", round(prom_quantitative_time$percentage), "%)", sep = ""))` cases, qualitatively stated (e.g., 'timely') in `r print(paste(prom_qualitative_time$count, "(", round(prom_qualitative_time$percentage), "%)", sep = ""))` cases, and not stated in `r print(paste(prom_notStated_time$count, "(", round(prom_notStated_time$percentage), "%)", sep = ""))` cases. Number of references allowed was quantitatively stated (e.g., 'max 20') in `r print(paste(prom_quantitative_ref$count, "(", round(prom_quantitative_ref$percentage), "%)", sep = ""))` cases and not stated in `r print(paste(prom_notStated_ref$count, "(", round(prom_notStated_ref$percentage), "%)", sep = ""))`. Whether critiques were sent for external peer-review was only stated in `r print(paste((prom_yes_review$count + prom_no_review$count), "(", round(prom_yes_review$percentage + prom_no_review$percentage), "%)", sep = ""))` cases, in `r print(paste(prom_yes_review$count, "(", round(prom_yes_review$percentage), "%)", sep = ""))` cases this was in the affirmative. 
    When stated quantitatively, length limits ranged from `r prom_length_range %>% pull(prom_min_length)` - `r prom_length_range %>% pull(prom_max_length)` words, time-to-submit limits ranged from `r prom_time_range %>% pull(prom_min_time)` -`r prom_time_range %>% pull(prom_max_time)` weeks, reference limits ranged between `r prom_ref_range %>% pull(prom_min_ref)` -`r prom_ref_range %>% pull(prom_max_ref)`. Table *XX* shows descriptive statistics for stated quantitative restrictions, broken down by critique type.

@@

# Create the second graph
graph3 <- d_journals %>%
  filter(sample_id == "random" | sample_id == "both") %>%
  filter(ppc_type == "A") %>%
  filter(!is.na(has_ppc)) %>%
  mutate(has_ppc = case_when(
    has_ppc == 'ARCHIVE' ~ 'Yes Implicit',
    has_ppc == 'NO EXPLICIT' ~ 'No Explicit',
    has_ppc == 'NO IMPLICIT' ~ 'No Implicit',
    has_ppc == 'YES' ~ 'Yes Explicit',
    TRUE ~ has_ppc
  )) %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = "", y = count, fill = has_ppc)) +
  geom_bar(stat = "identity",
           colour = "black") +
  theme_bw() +
  scale_fill_brewer(palette = "Oranges") +
  labs(x = "Journal Policy", y = "Count", fill = "Accepts PPC")

# Arrange both graphs in the same figure
combined_plots <- graph3 + graph4
combined_plots
```

``` {r echo = FALSE, message = FALSE}

# Calculate Wilson Confidence Intervals
sample_size <- 100
confidence_level <- 0.95

#CIs for 'Yes Explicit'
randomYes <- dj_random %>%
  filter(ppc_type == "A") %>%
  filter(has_ppc == "YES") %>%
  count()

ci_yesE <- binom.confint(randomYes$n, sample_size, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_yesE_LL <- round(ci_yesE$lower*100, 1)
ci_yesE_UL <- round(ci_yesE$upper*100, 1)

#CIs for 'Yes Implicit'
randomArchive <- dj_random %>%
  filter(ppc_type == "A") %>%
  filter(has_ppc == "ARCHIVE") %>%
  count()

ci_yesI <- binom.confint(randomArchive$n, sample_size, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_yesI_LL <- round(ci_yesI$lower*100, 1)
ci_yesI_UL <- round(ci_yesI$upper*100, 1)

#CIs for 'No Implicit'
randomNoImp <- dj_random %>%
  filter(ppc_type == "A") %>%
  filter(has_ppc == "NO IMPLICIT") %>%
  count()

ci_noI <- binom.confint(randomNoImp$n, sample_size, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_noI_LL <- round(ci_noI$lower*100, 1)
ci_noI_UL <- round(ci_noI$upper*100, 1)

#CIs for 'No Explicit'
randomNoExp <- dj_random %>%
  filter(ppc_type == "A") %>%
  filter(has_ppc == "NO EXPLICIT") %>%
  count()

ci_noE <- binom.confint(randomNoExp$n, sample_size, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_noE_LL <- round(ci_noE$lower*100, 1)
ci_noE_UL <- round(ci_noE$upper*100, 1)

# CIs for 2 types
random2types <- dj_random %>%
  filter(ppc_type == "B") %>%
  filter(!is.na(ppc_name)) %>%
  count()

ci_2typ <- binom.confint(random2types$n, sample_size, method = "wilson", conf.level = confidence_level)

# Convert the lower and upper limits to percentages and round them to one decimal place
ci_2typ_LL <- round(ci_2typ$lower*100, 1)
ci_2typ_UL <- round(ci_2typ$upper*100, 1)
```

@@

```
  As shown in Figure *XX*, post-publication critique policies were explicitly stated by `r dj_random %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "YES") / n() * 100, "%"))` of journals (`r paste0(round(confidence_level*100), "% CI [", ci_yesE_LL, "%, ", ci_yesE_UL, "%]")`), and we inferred that they were implicitly accepted (as denoted by archival examples) in `r dj_random %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "ARCHIVE") / n() * 100, "%"))` of journals (`r paste0(round(confidence_level*100), "% CI [", ci_yesI_LL, "%, ", ci_yesI_UL, "%]")`). `r dj_random %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "NO IMPLICIT") / n() * 100, "%"))` of journals (`r paste0(round(confidence_level*100), "% CI [", ci_noI_LL, "%, ", ci_noI_UL, "%]")`) had no advertised post-publication critique policy, and `r dj_random %>% filter(ppc_type == "A") %>% summarise(percentage = paste0(sum(has_ppc == "NO EXPLICIT") / n() * 100, "%"))` of journals (`r paste0(round(confidence_level*100), "% CI [", ci_noE_LL, "%, ", ci_noE_UL, "%]")`) explicitly stated that they did not accept critiques (*Clinical Psychology: Science and Practice and Psychological Bulletin*, also included in sample for prominent journals). `r paste0(random2types, "%")` of journals accepted two types (`r paste0(round(confidence_level*100), "% CI [", ci_2typ_LL, "%, ", ci_2typ_UL, "%]")`); *Cortex* offered commentaries and 'verification reports' (Chambers, 2020) and *Journal of Cognition and Culture* offered commentaries and letters.
